---
title: "COVID CO2 Emissions"
author: "Gavin McNicol"
output: html_document
---

## Introduction

How much carbon dioxide (CO2) do different countries emit, how have their emissions changed over time, and how were they impacted by the COVID-related restrictions on travel and economic activity during 2020?
Answering these questions (at a high level) is the focus of this analysis.

### Packages

We will use the R **tidyverse**, **lubridate**, and **scales** packages for data wrangling and visualization, and the **DT** package for interactive display of tabular output, and the **unvotes** package for the data.

*Packages are pre-assembled collections of functions and objects (often data)*

```{r load-packages, warning=FALSE, message=FALSE}
library(tidyverse)
library(lubridate)
library(scales)
library(DT)
```

### Data

The data we're using originally came from the (2020 Global Carbon Budget)[<https://doi.org/10.5194/essd-12-3269-2020>] which is a product of the (Global Carbon Project)[<https://doi.org/10.18160/gcp-2020>].
The data comes in an Excel file called `National_Carbon_Emissions_2020v1.0.xlsx` and has already been added to this exercise folder.
In the code chunk below we modify the data by joining the various data frames provided in the package to help you get started with the analysis.

```{r}
unvotes <- un_votes %>%
  mutate(country =
           case_when(
             country == "United Kingdom of Great Britain and Northern Ireland" ~ "UK & NI",
             country == "United States of America"                             ~ "US",
             TRUE                                                              ~ country
           )) %>%
  inner_join(un_roll_calls, by = "rcid") %>%
  inner_join(un_roll_call_issues, by = "rcid")
```

## UN voting patterns

Let's create a data visualisation that displays how the voting record of the UK & NI changed over time on a variety of issues, and compares it to two other countries: US and Turkey.

We can easily change which countries are being plotted by changing which countries the code above `filter`s for.
Note that the country name should be spelled and capitalized exactly the same way as it appears in the data.
See the [Appendix](#appendix) for a list of the countries in the data.

```{r plot-yearly-yes-issue, fig.width=10, fig.height=6, message=FALSE}
unvotes %>%
  filter(country %in% c("UK & NI", "US", "Turkey")) %>%
  mutate(year = year(date)) %>%
  group_by(country, year, issue) %>%
  summarize(percent_yes = mean(vote == "yes")) %>%
  ggplot(mapping = aes(x = year, y = percent_yes, color = country)) +
  geom_point(alpha = 0.4) +
  geom_smooth(method = "loess", se = FALSE) +
  facet_wrap(~issue) +
  scale_y_continuous(labels = percent) +
  labs(
    title = "Percentage of 'Yes' votes in the UN General Assembly",
    subtitle = "1946 to 2015",
    y = "% Yes",
    x = "Year",
    color = "Country"
  )
```

## References

1.  David Robinson (2017). [unvotes](https://CRAN.R-project.org/package=unvotes): United Nations General Assembly Voting Data. R package version 0.2.0.
2.  Erik Voeten "Data and Analyses of Voting in the UN General Assembly" Routledge Handbook of International Organization, edited by Bob Reinalda (published May 27, 2013).
3.  Much of the analysis has been modeled on the examples presented in the [unvotes package vignette](https://cran.r-project.org/web/packages/unvotes/vignettes/unvotes.html).

## Appendix {#appendix}

Below is a list of countries in the dataset:

```{r list-countries, echo=FALSE}
unvotes %>% 
  select(country) %>%
  arrange(country) %>% 
  distinct() %>%
  datatable()
```
