---
title: "Unit 1 - 3 Starter"
author: "TEST"
date: "TEST"
output: html_document
---

### Load packages and data

```{r load-packages, message=FALSE}
library(tidyverse) 
```

```{r load-data, message=FALSE}
sand_minerals <- read_csv("data/sand-minerals.csv")
site_descriptions <- read_csv("data/site-descriptions.csv")
density <- read_csv("data/density.csv")
solubility_hardness <- read_csv("data/solubility-hardness.csv")
```

## Exercises

### Exercise 1

```{r inspect-sand-minerals}
glimpse(sand_minerals)
names(sand_minerals)
```

There are 23 rows and 2 columns (variables).
Each row is a sand grain mineral observed at a particular location (one of ten).

### Exercise 2

```{r inspect-other-dataframes}
glimpse(density)
glimpse(site_descriptions)
glimpse(solubility_hardness)
names(density)
names(site_descriptions)
names(solubility_hardness)
```

density has 3 rows and 2 columns, site_descriptions has 10 rows and 2 columns, solubility_hardness has 3 rows and 3 columns

In `density` each row is the density of a particular mineral In `site_descriptions` each row is information about the location of a site In `solubility_hardness` each row is the solubility and hardness for a particular mineral

### Exercise 3

Remove this text, and add your answer for Exercise 3 here.
Add code chunks as needed.
Don't forget to label your code chunk.
Do not use spaces in code chunk labels.

### Exercise 4

Remove this text, and add your answer for Exercise 4 here.
Add code chunks as needed.
Don't forget to label your code chunk.
Do not use spaces in code chunk labels.

### Exercise 5

Remove this text, and add your answer for Exercise 5 here.
Add code chunks as needed.
Don't forget to label your code chunk.
Do not use spaces in code chunk labels.

### Exercise 6

Remove this text, and add your answer for Exercise 6 here.
Add code chunks as needed.
Don't forget to label your code chunk.
Do not use spaces in code chunk labels.
