---
title: "Weather Check"
author: "Gavin McNicol"
date: "`r Sys.Date()`"
output:
  html_document: 
    toc: yes
    toc_float: yes
---

In this mini analysis we work with the data used in the FiveThirtyEight story titled ["Where People Go to Check the Weather"](https://fivethirtyeight.com/features/weather-forecast-news-app-habits/).
Your task is to fill in the blanks denoted by `___`.

## Data and packages

We start with loading the packages we'll use.

```{r load-packages, message=FALSE}
library(fivethirtyeight)
library(tidyverse)
```

The dataset contains internet survey results on where `r nrow(weather_check)` people go to check the weather forecast.
Only `r sum(weather_check$ck_weather == TRUE)/nrow(weather_check)*100`% of respondents were checked the weather daily.

```{r}
weather_check_daily <- weather_check %>% 
  filter(ck_weather == TRUE)
```

There are `___` respondents who check the weather daily.

The `weather_check` variables we'll focus on are the following:

-   `weather_source`: Budget in 2013 inflation adjusted dollars
-   `age`: Domestic gross (US) in 2013 inflation adjusted dollars
-   `hhold_income`: Total International (i.e., worldwide) gross in 2013 inflation adjusted dollars

And we'll also use the `region` and `female` variable for **grouping**.

## Analysis

Let's take a look at how median budget and gross vary by whether the movie passed the Bechdel test, which is stored in the `binary` variable.

```{r message = FALSE}
bechdel90_13 %>%
  group_by(binary) %>%
  summarise(med_budget = median(budget_2013),
            med_domgross = median(domgross_2013, na.rm = TRUE),
            med_intgross = median(intgross_2013, na.rm = TRUE))
```

Next, let's take a look at how median budget and gross vary by a more detailed indicator of the Bechdel test result.
This information is stored in the `clean_test` variable, which takes on the following values:

-   `ok` = passes test
-   `dubious`
-   `men` = women only talk about men
-   `notalk` = women don't talk to each other
-   `nowomen` = fewer than two women

```{r message = FALSE}
bechdel90_13 %>%
  #group_by(___) %>%
  summarise(med_budget = median(budget_2013),
            med_domgross = median(domgross_2013, na.rm = TRUE),
            med_intgross = median(intgross_2013, na.rm = TRUE))
```

In order to evaluate how return on investment varies among movies that pass and fail the Bechdel test, we'll first create a new variable called `roi` as the ratio of the gross to budget.

```{r}
bechdel90_13 <- bechdel90_13 %>%
  mutate(roi = (intgross_2013 + domgross_2013) / budget_2013)
```

Let's see which movies have the highest return on investment.

```{r}
bechdel90_13 %>%
  arrange(desc(roi)) %>% 
  select(title, roi, year)
```

Below is a visualization of the return on investment by test result, however it's difficult to see the distributions due to a few extreme observations.

```{r warning = FALSE}
ggplot(data = bechdel90_13, 
       mapping = aes(x = clean_test, y = roi, color = binary)) +
  geom_boxplot() +
  labs(title = "Return on investment vs. Bechdel test result",
       x = "Detailed Bechdel result",
       y = "___",
       color = "Binary Bechdel result")
```

What are those movies with *very* high returns on investment?

```{r}
bechdel90_13 %>%
  filter(roi > 400) %>%
  select(title, budget_2013, domgross_2013, year)
```

Zooming in on the movies with `roi < ___` provides a better view of how the medians across the categories compare:

```{r warning = FALSE}
ggplot(data = bechdel90_13, mapping = aes(x = clean_test, y = roi, color = binary)) +
  geom_boxplot() +
  labs(title = "Return on investment vs. Bechdel test result",
       subtitle = "___", # Something about zooming in to a certain level
       x = "Detailed Bechdel result",
       y = "Return on investment",
       color = "Binary Bechdel result") +
  coord_cartesian(ylim = c(0, 15))
```
