---
title: "Lab 03 - Great Lakes Fish Stocking"
author: "Insert your name here"
date: "Insert date here"
output: html_document
---

### Load packages and data

```{r load-packages, message=FALSE}
library(tidyverse) 
```

```{r load-data, message=FALSE}
fish_stock <- read_csv("data/great-lakes-fish-stock.csv")
```

## Exercises

### Exercise 1

There are `r nrow(fish_stock)` observations and `r ncol(fish_stock)` variables in this dataset.
Each row represents a fish stocking event in one of the Great Lakes.

### Exercise 2

```{r filter-stock-data, echo = FALSE, eval = TRUE}
salmon_stock <- fish_stock %>%
  filter(
    SPECIES %in% c("ATS", "CHS", "COS"),
    !is.na(WEIGHT),
     LAKE != "ON"
  )
```

```{r summarize-new-data}
glimpse(salmon_stock)
```

The new data frame `salmon_stock` contains 5,131 rows, 8 columns (variables).
YEAR, MONTH, NO_STOCKED, AGEMONTH and WEIGHT are numerical.
LAKE, STATE_PROV, SPECIES are character strings.

### Exercise 3

```{r compute-average-weight, eval = TRUE}
salmon_stock <- salmon_stock %>%
  mutate(
    AVG_WEIGHT = WEIGHT / NO_STOCKED * 1000
  )
```

The mutate function creates a new variable `AVG_WEIGHT` which is the average weight of the fish stocked, in grams, and is computed as NO_STOCKED divided by WEIGHT multiplied by 1000.

### Exercise 4

```{r eval = TRUE}
salmon_stock %>% 
  group_by(LAKE) %>% 
  filter(YEAR > 2000) %>% 
  summarize(MEAN_NO_STOCKED = mean(NO_STOCKED, na.rm = T),
            MEAN_WEIGHT = mean(WEIGHT, na.rm = T)) %>% 
  arrange(desc(MEAN_NO_STOCKED))
```

Lake Huron had the largest mean number of salmon stocked per event, and Lake Michigan has the largest mean weight of salmon stocked per event.

### Exercise 5

...

### Exercise 6

...
