---
title: "EAES 494 Midterm Instructions"
output: 
  tufte::tufte_html:
    css: ../exam.css
    tufte_variant: "envisioned"
    highlight: pygments
link-citations: yes
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(eval = TRUE)
```

# ---- Exam Start ----

## Academic Honesty Statement

The following statement is provided in your `midterm-starter.Rmd`:

*I, \_\_\_\_\_\_\_\_\_\_\_\_\_, hereby state that I will not communicate with or gain information in any way from my classmates, or copy work from past assignments, during this exam, and that all work is my own.*

Please complete the blank (`___`) with your name.

## Warm up

Before we start, let's warm up by updating the YAML.

-   Change the author name and date in the starter file YAML.
-   **Knit** the document. 🧶
-   **Commit** your changes with a meaningful commit message. ✅
-   **Push** your changes to GitHub. ⬆️

## Loading packages and Data

We will need `{tidyverse}` packages and functions, and will be using the `penguins` dataset from the `{palmerpenguins}` package.
The code chunk below will load the two packages and is also provided in your starter file.

```{r load-packages, message = F}
library(tidyverse)
library(palmerpenguins)
```

# Inspecting Data

There are several ways to inspect a data frame.
One method is to use the `glimpse()` function.
Other methods include using `View()` or `?` help pages.

1.  **(5 points)** Under Ex. 1 in your starter file, write and label a code chunk that uses `glimpse()` to allow you to inspect the `penguins` dataset. After your code chunk, state **(1)** how many rows and variables there are in the `penguins` dataset and **(2)**, of the variables, how many are categorical and how many are numerical.

In a tidy data frame, each column is a variable, and each row is an observation.

2.  **(5 points)** Under Ex. 2 in your starter file, write and label a code chunk that uses `names()` to allow you to inspect the variable names in the `penguins` dataset. After your code chunk, state two stylistic choices for these variable names that means they accord with good R syntax.

Each row of the `penguins` dataset describes the physical characteristics of a penguin sampled in Antarctica.
We might wonder: how many observations do we have for each `species`?

3.  **(5 points)** Under Ex. 3 in your starter file, write and label a code chunk that uses the `count()` function to output the number of observations (rows) for each `species` of penguin in the `penguins` dataset. **Hint:** remember you can use help pages to look at function syntax. After your code chunk, state the pair of data wrangling functions we could have used instead to generate a similar output.

# Exploratory visualization

Rather than looking at a summary table of species counts, we might want to visualize the number of penguins of each species, observed for each island studied.

4.  **(5 points)** Under Ex. 4 in your starter file, copy and complete the code chunk provided below that will generate a bar plot of penguin observations by `island`. The fill aesthetic of each bar will vary with `species`. Under your completed code chunk, for each species of penguin, state which islands they were observed at.

```{r visualize-penguin-island-species, eval = F}
___ %>% 
  ggplot(aes(x = ___, fill = ___)) +
  geom_bar() +
  theme_classic()
```

Histograms also provide useful visualizations of data.
We might wonder: how does penguin body mass vary across the different species?

5.  **(5 points)** Under Ex. 5 in your starter file, copy and complete the code chunk provided below that will generate histograms of penguin `body_mass_g`. Three histogram facets will be output, one for each `species`. The fill aesthetic of each histogram will also vary with `species`. Under the completed code chunk, state whether you think the center (average) of body mass depends on species or not, and why.

```{r visualize-penguin-body-mass, eval = F}
___ %>% 
  ggplot(aes(x = ___, fill = ___)) +
  geom_histogram() +
  facet_wrap(~___, 
             ncol = 1) +
  theme_classic()
```

# Wrangling data

We might want to just focus on analyzing penguin flipper length and how it varies with the categorical variables.

6.  **(10 points)** Under Ex. 6 in your starter file, write and label a code chunk that takes `penguins` as input and uses a pipeline of wrangling functions to: (1) **select** `flipper_length_mm` and all the categorical variables; (2) **filter** for *non-NA* values in each variable (using `!is.na(variable_name)`); and (3) assign the output to a new data object called `penguins_flippers`.

# Recoding data

The following code chun0k, which is also provided for you in the starter file, uses a violin plot to visualize the distribution of penguin flipper lengths by-species, from our newly created `penguins_flippers` object.

```{r visualize-penguin-flipper-length, eval = F}
penguins_flippers %>% 
  ggplot(aes(x = sex, y = flipper_length_mm)) +
  geom_violin() +
  theme_classic()
```

7.  **(5 points)**

-   Under Ex. 7 in your starter file, state why `female` appears on the left and why `male` appears on the right.
-   In your response, make reference to the data type/class of the sex `variable` and the two components that make up that data class.
-   Then, copy the code chunk below into your starter file, and complete it such that it outputs a violin plot with `male` on the left and `female` on the right.

```{r recode-sex-levels, eval = F}
___ %>% 
  mutate(sex = fct_relevel(sex, "___", "___")) %>% 
    ggplot(aes(x = sex, y = flipper_length_mm)) +
    geom_violin() +
    theme_classic()
```

# Reading and correcting data

There is a version of `penguins` in the `/data` folder within your midterm project.
Take a look at what it is called.
There are a few issues with this version of the penguins dataset.
**All the problems with the data are located within the top 5 rows.**

8.  **(10 points)**

-   Write and label a code chunk that will read the problematic dataset from your data folder, into R, and assign the data to the object `penguins_with_issues`.
-   In a second, separate *line of code*, under your read function, use `glimpse()` to view the columns and first few rows of `penguins_with_issues`.
-   Identify and state under your code chunk, the two types of problems with this data set.
-   Finally, copy and paste your first read code chunk, re-label it.
-   Edit the second code chunk so that the data is corrected during or after it is read in. (**Hint**: Examine the argument options in the read function help page.)

🧶 ✅ ⬆️ Knit, *commit, and push your final changes to GitHub with an appropriate commit message.*

# ---- Exam Finish ----
