---
title: "Homework 04 - Data Classes w. forcats & lubridate"
author: "RUBRIC"
date: "DATE"
output: html_document
---

### Load packages and data

```{r load-packages, message=FALSE}
library(tidyverse)
library(openintro) 
library(lubridate)
```

## Exercises

### Exercise 1

```{r create-3-vectors}
numeric <- c(1:5)
character <- c("a", "b", "c", "d", "e")
mixed <- c(1, "b", 3, "d", 5)
```

-   numeric is a `r typeof(numeric)` vector of class `r class(numeric)`
-   character is a `r typeof(character)` vector of class `r class(character)`
-   mixed is a `r typeof(mixed)` vector of class `r class(mixed)`

### Exercise 2

```{r create-data-frame, eval = T}
data_frame <- bind_cols("numeric" = numeric, "character" = character, "mixed" = mixed)
data_frame
```

-   data_frame is of **type** `r typeof(data_frame)` and of **class** `r class(data_frame)`

### Exercise 3

```{r create-list}
numeric <- c(1:4)
character <- c("a", "b", "c", "d", "e")
mixed <- c(1, "b", 3, "d", 5)
# data_frame <- bind_cols("numeric" = numeric, "character" = character, "mixed" = mixed)
```

`bind_cols()` won't work because data frames need to be created from vectors of equal length.

### Exercise 4

```{r my-list, eval = T}
my_list <- list("numeric" = numeric, "character" = character, "mixed" = mixed)
my_list
```

the `list()` function can concatenate vectors of different length

### Exercise 5

Remove this text, and add your answer for Exercise 5 here.

### Exercise 6

Remove this text, and add your answer for Exercise 6 here.

### Exercise 7

Remove this text, and add your answer for Exercise 7 here.

### Exercise 8

Remove this text, and add your answer for Exercise 8 here.

### Exercise 9

Remove this text, and add your answer for Exercise 9 here.

### Exercise 10

Remove this text, and add your answer for Exercise 10 here.
