---
title: "Global CO2 Emissions"
author: "Gavin McNicol"
date: "`r Sys.Date()`"
output:
  html_document: 
    toc: yes
    toc_float: yes
---

## Introduction

How much carbon dioxide (CO2) do different countries emit, and how have their emissions changed over time?
Answering these questions (at a high level) is the focus of this analysis.

### Packages

We will use the R **tidyverse**, **lubridate**, and **scales** packages for data wrangling and visualization, and the **DT** package for interactive display of tabular output, and the **unvotes** package for the data.

*Packages are pre-assembled collections of functions and objects (often data)*

```{r load-packages, warning=FALSE, message=FALSE}
library(tidyverse)
library(lubridate)
library(scales)
library(DT)
```

### Data

The data we're using originally came from the [2020 Global Carbon Budget](https://doi.org/10.5194/essd-12-3269-2020) which is a product of the [Global Carbon Project](https://doi.org/10.18160/gcp-2020).
The data comes in an Excel file called `National_Carbon_Emissions_2020v1.0.xlsx` and has already been added to this exercise folder.

In the code chunk below:

-   the first line reads in the data, selects the correct tab `Territorial Emissions`, and removes 11 header rows.
-   the next line renames the first column `year`.
-   the third line rearranges the country emission data for easier plotting.
-   the final line converts the `emissions` data column to numerical values.

This will help you get started with the analysis.

```{r message = F}
emissions <- readxl::read_excel("National_Carbon_Emissions_2020v1.0.xlsx", sheet = "Territorial Emissions", skip = 11) %>% 
  rename("year" = `...1`) %>% 
  gather(key = "territory", value = "emissions", 2:234) %>% 
  mutate(emissions = as.numeric(emissions))
```

## Territory CO2 Emissions

Let's create a data visualization that displays how the CO2 emissions of the USA changed over time on a variety of issues, and compares it to three other countries: the United Kingdom, China, and Venezuela.

We can easily change which countries are being plotted by changing which countries the code above `filter`s for.
Note that the country name should be spelled and capitalized exactly the same way as it appears in the data.

See the [Appendix](#appendix) for a list of the countries in the data.

Next, let's make the plot:

```{r plot-yearly-co2-emissions, fig.width=5, fig.height=3, message=FALSE}
emissions %>%
  filter(territory %in% c("USA", "United Kingdom", "China") &
           year > 1979) %>%
  ggplot(mapping = aes(x = year, y = emissions, color = territory)) +
  geom_point(alpha = 0.4) +
  geom_smooth(method = "loess", se = FALSE) +
  labs(
    title = "Emissions of Carbon Dioxide by Territory",
    subtitle = "1980 to 2020",
    y = expression("Emissions (Mt C y"^{-1}*")"),
    x = "Year",
    color = "Territory"
  )
```

## Exercises

Edit the code in the chunk above to:

1.  Visualize an additional territory/territories of your choice.
2.  Shorten the x-axis to only show years after 2009.
3.  Update the plot subtitle to reflect the shorter time period.
4.  Double the size of the entire plot (clue: you may have to "think outside the box").
5.  Finally, edit the `YAML` (document metadata) so that your name shows instead of `Gavin McNicol`.

*Remember to `Knit` the file when you make a change. This will save your work and create an .html output file.*

## References

1.  Global Carbon Project.
    (2020).
    Supplemental data of Global Carbon Budget 2020 (Version 1.0) [Data set].
    [Global Carbon Project](https://doi.org/10.18160/gcp-2020)

2.  This exercise was modeled after the UN Votes application exercise in [Data Science in a Box](https://datasciencebox.org/index.html).

## Appendix {#appendix}

Below is a list of countries in the dataset:

```{r list-countries, echo=FALSE}
emissions %>% 
  dplyr::select(territory) %>% 
  distinct() %>% 
  datatable()
```
