---
title: "HW 02 - Water and Sanitation - Answer Key"
author: "Gavin McNicol"
date: "`r Sys.Date()`"
output: github_document
---

## Load packages and data

```{r load-packages-data, message = FALSE}
library(tidyverse)
wat_san <- read_csv("data/water-and-sanitation.csv")
```

## Exercises

### Exercise 1

There are `r nrow(wat_san)` observations in the dataset.

### Exercise 2

Each row in `wat_san` corresponds to a different year of water and sanitation access observations for each country/entity.

### Exercise 3

Rates of open defecation appear have been reduced.
There are fewer outlier countries with very high rates, and the height of the lowest possible, 0-1% rate (depends on student choice of bin size for upper limit here), has increased since 2000.

```{r no-sanitation-years}
ggplot(data = wat_san, mapping = aes(x = sanitation_none)) +
  geom_histogram(binwidth = 1) +
  facet_wrap(~Year)
```

### Exercise 4

It does not appear that the complete absence of water and sanitation access are well correlated in the 5 countries with the lowest sanitation access.

```{r bottom-5-2000}
wat_san %>% 
  filter(Year == 2000) %>% 
  slice_max(sanitation_none, n = 5) %>% 
  arrange(desc(sanitation_none))
```

### Exercise 5

This does support my previous conclusion that there is no clear relationship between `water_none` and `sanitation_none` because there is so much scatter in the data.

```{r all-2000-plot}
wat_san %>% 
  filter(Year == 2000) %>% 
  ggplot(aes(x = water_none, y = sanitation_none)) + 
  geom_point()
```

### Exercise 6

The countries with the largest percentage of the population with no sanitation access have changed between 2000 and 2020.
Ethiopia, Cambodia, Eritrea, and India are no longer in the bottom 5 but Chad, South Sudan, Benin and Namibia are now present.

```{r bottom-5-2020}
wat_san %>%
  filter(Year == 2020) %>% 
  slice_max(sanitation_none, n = 5) 
```

### Exercise 7

Between 2000 and 2020, the summary statistics all decreased in values, indicating lower amounts of open defecation and lack of access to clean water.

Overall, values for no sanitation are still higher than values for no water access.

```{r progress-summary}
wat_san %>% 
  filter(Year == 2000 & Entity != "World" | Year == 2020 & Entity != "World") %>% 
  group_by(Year) %>% 
  summarize(water_none_min = min(water_none, na.rm = T),
            water_none_mean = mean(water_none, na.rm = T),
            water_none_median = median(water_none, na.rm = T),
            water_none_stdev = sd(water_none, na.rm = T),
            water_none_iqr = IQR(water_none, na.rm = T),
            water_none_max = max(water_none, na.rm = T),
            
            sanitation_none_min = min(sanitation_none, na.rm = T),
            sanitation_none_mean = mean(sanitation_none, na.rm = T),
            sanitation_none_median = median(sanitation_none, na.rm = T),
            sanitation_none_stdev = sd(sanitation_none, na.rm = T),
            sanitation_none_iqr = IQR(sanitation_none, na.rm = T),
            sanitation_none_max = max(sanitation_none, na.rm = T)

            )
  
```

### Exercise 8

Cambodia and India both dropped out of the top 5 for open defecation because rates of open defecation dropped faster there than the world average, and much faster than in Chad and Niger, where rates stayed almost constant.

```{r country-global-changes}
wat_san %>% 
  filter(Entity == "World" | Entity == "India" | Entity == "Cambodia" | Entity == "Niger" | Entity == "Chad") %>% 
  ggplot(aes(Year, sanitation_none, color = Entity)) +
  geom_line()
```
