---
title: "Lab 03 - Great Lakes Fish Stocking"
output: 
  tufte::tufte_html:
    css: ../lab.css
    tufte_variant: "envisioned"
    highlight: pygments
link-citations: yes
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(eval = TRUE)
```

# Great Lakes Fish Stocking Database

For Lab 03 you'll be using the [Great Lakes Fish Stocking database](http://www.glfc.org/fishstocking/index.htm) which provides access to a centralized, comprehensive database of all fish stocked from artificial propagation into the Great Lakes since 1950.
The database was developed through a peer review process that involved fishery professionals from every agency that stocks fish into the Great Lakes.
Fishery managers use this database to help them understand the food and habitat needs of stocked predators in each lake.
The database is updated, maintained, and hosted by the U.S.
Fish and Wildlife Service.

# Learning goals

-   Practicing data wrangling and data visualization

# Getting started

If you're looking at these instructions you've already done the following steps, but as a reminder:

-   Go to the course GitHub organization and locate your Unit 3 repo, which should be named `unit-3-YOUR_LAST_NAME`.
-   Grab the URL of the repo, and clone it in RStudio.
-   First, open the R Markdown instructions `lab-O3-great-lakes-fish-stock.Rmd` and the lab starter `lab-03.Rmd` and Knit them
-   Make sure they compile without errors.

## Warm up

Before we introduce the data, let's warm up with some simple exercises.

-   Update the YAML, changing the author name to your name, and **knit** the document. 🧶
-   Commit your changes with a meaningful commit message. ✅
-   Push your changes to GitHub.️ ⬆️
-   Go to your repo on GitHub and confirm that your changes are visible in your Rmd files.

## Packages

We'll use the **tidyverse** package for much of the data wrangling.
This package is already installed for you.
You can load them by running the following in your Console:

```{r load-packages, message = FALSE}
library(tidyverse)
```

## Data

The dataset for this assignment can be found as a CSV (comma separated values) file in the `data` folder of your repository.
You can read it in using the following.

```{r load-data, message = FALSE}
fish_stock <- read_csv("data/great-lakes-fish-stock.csv")
```

The variable descriptions are as follows:

-   `YEAR`: year of stocking, YYYY

-   `MONTH`: month of stocking, not always provided by agency

-   `LAKE`: Great Lake, coded

-   

        ER (Lake Erie)

-   

        HU (Huron)

-   

        MI (Lake Michigan)

-   

        ON (Lake Ontario)

-   

        SC (Lake St. Clare)

-   

        SU (Lake Superior)

-   `STATE_PROVENCE`: state or province where stocking occurred

-   

        IL (Illinois)

-   

        IN (Indiana)

-   

        MI (Michigan)

-   

        MN (Minnesota)

-   

        NY (New York)

-   

        OH (Ohio)

-   

        ON (Ontario)

-   

        PA (Pennsylvania)

-   

        WI (Wisconsin)

-   `SPECIES`: species coded by naming convention (

-   `NO_STOCKED`: number of fish stocked

-   `AGEMONTH`: age in months, not always provided by agency

-   `WEIGHT`: total weight of fish stocked in kilograms

## Get to know your data

1.  How many observations and how many variables are in the dataset? Use inline code to answer this question. What does each row represent?

There are some observations in this dataset that we will exclude from our analysis to match the Buzzfeed results.

2.  Create a new data frame called `nobel_living` that filters for

-   laureates for whom `country` is available
-   laureates who are people as opposed to organizations (organizations are denoted with `"org"` as their `gender`)
-   laureates who are still alive (their `died_date` is `NA`)

```{r echo=FALSE}
nobel_living <- nobel %>%
  filter(
    !is.na(country),
    gender != "org",
    is.na(died_date)
  )
```

Confirm that once you have filtered for these characteristics you are left with a data frame with `r nrow(nobel_living)` observations, once again using inline code.

🧶 ✅ ⬆️ Knit, *commit, and push your changes to GitHub with an appropriate commit message. Make sure to commit and push all changed files so that your Git pane is cleared up afterwards.*

## Most living Nobel laureates were based in the US when they won their prizes

... says the Buzzfeed article.
Let's see if that's true.

First, we'll create a new variable to identify whether the laureate was in the US when they won their prize.
We'll use the `mutate()` function for this.
The following pipeline mutates the `nobel_living` data frame by adding a new variable called `country_us`.
We use an if statement to create this variable.
The first argument in the `if_else()` function we're using to write this if statement is the condition we're testing for.
If `country` is equal to `"USA"`, we set `country_us` to `"USA"`.
If not, we set the `country_us` to `"Other"`.

```{marginfigure}
Note that we can achieve the same result using the `fct_other()` function we've seen before (i.e. with `country_us = fct_other(country, "USA")`). We decided to use the `if_else()` here to show you one example of an if statement in R.
```

```{r}
nobel_living <- nobel_living %>%
  mutate(
    country_us = if_else(country == "USA", "USA", "Other")
  )
```

Next, we will limit our analysis to only the following categories: Physics, Medicine, Chemistry, and Economics.

```{r}
nobel_living_science <- nobel_living %>%
  filter(category %in% c("Physics", "Medicine", "Chemistry", "Economics"))
```

For the next exercise work with the `nobel_living_science` data frame you created above.
This means you'll need to define this data frame in your R Markdown document, even though the next exercise doesn't explicitly ask you to do so.

3.  Create a faceted bar plot visualizing the relationship between the category of prize and whether the laureate was in the US when they won the nobel prize.
    Interpret your visualization, and say a few words about whether the Buzzfeed headline is supported by the data.

    -   Your visualization should be faceted by category.
    -   For each facet you should have two bars, one for winners in the US and one for Other.
    -   Flip the coordinates so the bars are horizontal, not vertical.

🧶 ✅ ⬆️ Knit, *commit, and push your changes to GitHub with an appropriate commit message. Make sure to commit and push all changed files so that your Git pane is cleared up afterwards.d*

## But of those US-based Nobel laureates, many were born in other countries

```{marginfigure}
**Hint:** You should be able to ~~cheat~~ borrow from code you used earlier to create the `country_us` variable.
```

4.  Create a new variable called `born_country_us` that has the value `"USA"` if the laureate is born in the US, and `"Other"` otherwise.
    How many of the winners are born in the US?

5.  Add a second variable to your visualization from Exercise 3 based on whether the laureate was born in the US or not.
    Based on your visualization, do the data appear to support Buzzfeed's claim?
    Explain your reasoning in 1-2 sentences.

    -   Your final visualization should contain a facet for each category.
    -   Within each facet, there should be a bar for whether the laureate won the award in the US or not.
    -   Each bar should have segments for whether the laureate was born in the US or not.

🧶 ✅ ⬆️ Knit, *commit, and push your changes to GitHub with an appropriate commit message. Make sure to commit and push all changed files so that your Git pane is cleared up afterwards.*

## Here's where those immigrant Nobelists were born

```{marginfigure}
Note that your bar plot won't exactly match the one from the Buzzfeed article. This is likely because the data has been updated since the article was published.
```

6.  In a single pipeline, filter for laureates who won their prize in the US, but were born outside of the US, and then create a frequency table (with the `count()` function) for their birth country (`born_country`) and arrange the resulting data frame in descending order of number of observations for each country. Which country is the most common?

🧶 ✅ ⬆️ Knit, *commit, and push your changes to GitHub with an appropriate commit message. Make sure to commit and push all changed files so that your Git pane is cleared up afterwards and review the md document on GitHub to make sure you're happy with the final state of your work.*

Now go back through your write up to make sure you've answered all questions and all of your R chunks are properly labelled.
