---
title: "Lab 04 - Data Type Coercion"
author: "RUBRIC"
date: "DATE"
output: html_document
---

**Half point for YAML MAX (0.5)**

### Load packages and data

```{r load-packages, message=FALSE}
library(openintro) 
```

## Exercises

### Exercise 1

**Half point per chunk MAX (1.5)**

```{r}
1
1L
"C"
```

-   `1` is type `r typeof(1)`
-   `1L` is type `r typeof(1L)`
-   `"C"` is type `r typeof("C")`

```{r}
c(1, 1L, "C")
```

The first chunk prints out each value as a separate object, second code concatenates the objects as elements of a single vector.

```{r}
typeof(c(1, 1L, "C"))
```

The vector type is character.
R automatically coerces numeric to character when both types are present in a vector.

### Exercise 2

**Half point for each narrative (MAX 1)**

```{r}
1:3
5
```

-   `1:3` is type `r typeof(1:3)`
-   `5` is type `r typeof(5)`

```{r}
typeof(c(1:3, 5))
```

The vector type is double.
R automatically coerces numeric integers to numeric doubles when both types are present in a vector.

### Exercise 3

**Half point for each narrative (MAX 1)**

```{r}
3 
FALSE
T
```

-   `3` is type `r typeof(3)`
-   `FALSE` is type `r typeof(FALSE)`
-   `T` is type `r typeof(T)`

```{r}
typeof(c(3, FALSE, T))
```

The vector is double.
R coerces logical data to numeric doubles when both are present.

### Exercise 4

**Half point for each narrative (MAX 1)**

```{r}
TRUE
as.numeric(TRUE)
```

-   `TRUE` is type `r typeof(TRUE)`
-   `as.numeric(TRUE)` is type `r typeof(as.numeric(TRUE))`

```{r}
typeof(c(TRUE, as.numeric(TRUE)))
```

Vector type is double.
R coerces logical to numeric when both are present in a vector.

### Exercise 5

**Half point for each narrative (MAX 1)**

```{r}
NA
FALSE
as.logical(1)
```

-   `NA` is type `r typeof(NA)`
-   `FALSE` is type `r typeof(FALSE)`
-   `as.logical(1)` is type `r typeof(as.logical(1))`

```{r}
typeof(c(NA, FALSE, as.logical(1)))
```

Vector is type logical.
All the terms are already logical.
R does not need to coerce any element of the vector.

### Exercise 6

**Half point for each narrative (MAX 1)**

```{r}
1L
0
1L / 0
"A"
```

-   `1L` is type `r typeof(1L)`
-   `0` is type `r typeof(0)`
-   `1L / 0` is type `r typeof(1L / 0)`
-   `"A"` is type `r typeof("A")`

```{r}
typeof(c(1L / 0, "A"))
```

Vector is type character.
Vectors are coerced to character type when character and numeric are present.
